---
title: "CaseStudy1_Group3"
author: "Limin Zheng"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Uncover findings from beers and breweries datasets.
1.	How many breweries are present in each state?
2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
3.	Report the number of NA's in each column.
4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
6.	Summary statistics for the ABV variable.
7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r echo=TRUE}
#load beers and breweries data into r. 
beers <- read.csv("Beers.csv", header = TRUE, sep = ",")
breweries <- read.csv("Breweries.csv", header=TRUE, sep = ",")
#check how many breweries are present in each state
library(dplyr)
brewperstate <- breweries %>% group_by(State) %>% summarize(n=n())
head(brewperstate)
```

```{r}
#outer merge the beers data with breweries data.
library(tidyr)
combine = merge(beers,breweries, by.x = "Brewery_id", by.y = "Brew_ID", all = TRUE)
#check the first 6 rows of combined data.
head(combine)
#check the last 6 rows of combined data.
tail(combine)
```

```{r echo=TRUE}
#Report the number of NA's in each column.
na.col <- colSums(is.na(combine))
na.col
#drop the rows with NA.
combine.clean = drop_na(combine)
write.csv(combine.clean, "combine.clean.csv")
```

```{r echo=TRUE}
#Get the median alcohol content(ABV) for each state. 
ABV <- aggregate(ABV~State, data=combine.clean, median)
#Get the median international bitterness unit(IBU) for each state.
IBU <- aggregate(IBU~State, data=combine.clean, median)
#combine the ABV, IBU data into a dataframe by states
ABV.IBU <- merge(ABV,IBU, by = "State", all = TRUE)
#Plot a bar chart to compare.
library(ggplot2)
ggplot(ABV.IBU, aes(x=reorder(State, -ABV),y=ABV, fill=IBU))+labs(title="Plot of ABV by state", x="State", y="ABV")+geom_bar(stat="identity") + theme_bw() +theme(axis.text.x=element_text(angle=90, size = 7)) +theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#get the state of maximun ABV
combine.clean[which(combine.clean$ABV == max(combine.clean$ABV)), "State"]
#get the state of maximum IBU
combine.clean[which(combine.clean$IBU == max(combine.clean$IBU)), "State"]
```

```{r}
summary(combine$ABV)
```

```{r}
ggplot(combine.clean, aes(ABV, IBU, color = State)) +
  geom_point(shape = 16, size = 3, show.legend = FALSE) +
  theme_minimal()
```

